Architectures for workflow integration in science gateways

### 2016-04-20 paper.tex
updates in section 2.1 and 2.2

At the same time, specific toolboxes were emerging in various
scientific domains to facilitate the assembly of software components
in consistent pipelines. For instance, in neuroimaging, our primary
domain of interest, tools such as Nipype~\cite{gorgolewski2011nipype},
PSOM~\cite{bellec2012pipeline}, SPM~\cite{ashburner2011spm} and
FSL~\cite{Jenkinson2012782} provide abstractions and functions to
handle the data and computing flow between processes implemented in a
variety of programming languages. Such tools were interfaced to
computing systems, in particular clusters, to create tasks, handle
their dependencies, and execute them. For instance, FSL can launch
tasks on SGE through its \texttt{fsl\_sub} tool, Nipype has execution
plugins for S/OGE, Torque, IPython, ssh and multi-processor servers,
and PSOM \todo{...}. Some of these tools also support advanced
features such as provenance tracking (Nipype, PSOM), redundancy
detection across analyses to avoid re-computation (PSOM), and so on. A
wide-array of workflows have been implemented using these pipelines
and are now widely shared across neuroimaging groups
world-wide. \todo{Provide more context on that?} These domain-specific
engines represent a tremendous opportunity for science gateways to
leverage existing tools and applications. They nicely complement
e-Science engines that are more oriented towards the exploitation of
distributed computing infrastructures, in particular grids and clouds.

@@

Science Gateway: An ecosystem, usually accessible through a web
  portal, that provides tools to access distributed
  infrastructures. Tools usually help users manage data transfers,
  task execution and authentication on multiple computing and storage
  locations. Examples: CBRAIN, VIP, NSG, etc

>>>>

At the same time, specific toolboxes were emerging in various
scientific domains to facilitate the assembly of software components
in consistent pipelines. For instance, in neuroimaging, our primary
domain of interest, tools such as Nipype~\cite{gorgolewski2011nipype},
PSOM~\cite{bellec2012pipeline}, SPM~\cite{ashburner2011spm} and
FSL~\cite{Jenkinson2012782} provide abstractions and functions to
handle the data and computing flow between processes implemented in a
variety of programming languages. Such tools were interfaced to
computing systems, in particular clusters, to create tasks, handle
their dependencies, and execute them. For instance, FSL can launch
tasks on SGE through its \texttt{fsl\_sub} tool, Nipype has execution
plugins for S/OGE, Torque, IPython, ssh and multi-processor servers,
and PSOM is interfaced to Torque, SGE, MOAB, LSF and Condor. Some of
these tools also support advanced features such as provenance tracking
(Nipype, PSOM), redundancy detection across analyses to avoid
re-computation (PSOM), and so on. A wide-array of workflows have been
implemented using these pipelines and are now widely shared across
neuroimaging groups world-wide. \todo{Provide more context on that?}
These domain-specific engines represent a tremendous opportunity for
science gateways to leverage existing tools and applications. They
nicely complement e-Science engines that are more oriented towards the
exploitation of distributed computing infrastructures, in particular
grids and clouds.

@@

Science gateways are ecosystems of software tools and services to
access distributed infrastructures, usually through a web interface.
Tools usually help users manage data transfers, task execution and
authentication on multiple computing and storage locations. Workflow
engines are of course part of this ecosystem since they are useful
systems to implement and execute applications.

Several models

Examples: CBRAIN, VIP, NSG, etc
